### 2025-05-01 2_Slides/SlidesFinal.qmd
Reorganized repo

---
title: "Crime in Los Angeles: Homicide Increases in the COVID-19 Era"
format:
  revealjs:
    theme: solarized
    slide-number: true
    chalkboard:
      buttons: false
    preview-links: auto
    footer: Group 5 - STAT 405 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Crime in Los Angeles
    title-slide-attributes:
      data-background: "#0f172a"
      data-transition: fade
server: shiny
---

# Introduction {background="#6B6B6B"}

## Why research crime in Los Angeles?

::: incremental
-   Homicide rates in Los Angeles increased substantially during pandemic
-   One of the largest increases in the country, alongside Houston and Philadelphia
-   Caused by pre-pandemic and pandemic-induced economic instability and inequalities
-   Los Angeles makes their crime data publicly available (2010-Present)
:::

::: footer
Learn more: [Thurgood Marshall Institute Report](https://www.naacpldf.org/the-truth-behind-crime-statistics/)
:::

## Our hypothesis

We predicted that:

::: incremental
-   Socioeconomic conditions have improved in Los Angeles since the pandemic

-   As these socioeconomic measures have bounced back, the rate of homicide has returned to a more "normal" rate

-   A similar trend might be seen with overall crime rates in LA

-   We also examined if increased police funding is correlated with a reduction in the number of homicides.
:::

## Primary Data

Our primary datasets are the "Crime Data from 2010 to 2019" and "Crime Data from 2020 to Present" National Incident-Based Reporting System (NIBRS) datasets, available on [Data.gov](https://catalog.data.gov/).

```{r, echo=FALSE}
library(ggplot2)
library(RSQLite)
library(dplyr)
con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))

crime_la_annually <- dbGetQuery(con, "SELECT * FROM crime_la_annually") %>% group_by(year) %>% summarize(obs = first(total_crimes))

dbDisconnect(con)

ggplot(crime_la_annually, aes(x = year, y = obs/1000)) +
  geom_col(fill = "#19297C") +
  ggthemes::theme_fivethirtyeight() +
  labs(title = "Number of Annual Observations in Primary Dataset",
       y = "Thousands of Observations (Crimes)",
       x = "Year") +
  theme(axis.title = element_text())
```

## Secondary

Our supplementary datasets were:

-   [Los Angeles County Eviction Data](https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/)

-   [City of Los Angeles Eviction Data](https://housing.lacity.gov/residents/renters/eviction-notices-filed)

-   [Los Angeles County COVID Data](http://dashboard.publichealth.lacounty.gov/covid19_surveillance_dashboard/)

-   [Census Bureau Median Income Data](https://data.census.gov/table/ACSST1Y2023.S1903?q=income)

-   [Los Angeles Police Department Budget Data](https://openbudget.lacity.org/#!/year/2023/operating/0/department_name/Police/0/program_name?vis=lineChart)

-   [Los Angeles Area Unemployment Rate Data](https://www.bls.gov/regions/west/ca_losangeles_msa.htm#eag)

# Data Exploration {background="#6B6B6B"}

## Weekly Crime Counts in Los Angeles from 2010 to present

```{r}
#| fig.align: 'center'
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(
  con,
  "
  SELECT
    YearWeek,
    MIN(Date) AS start_of_week,
    MAX(Date) AS end_of_week,
    COUNT(*)  AS Crimes
  FROM crime_la
  GROUP BY YearWeek
  ORDER BY YearWeek
  "
)
weekly_counts <- dbFetch(query, n = -1)
dbClearResult(query)
dbDisconnect(con)

weekly_counts$start_of_week <- as.Date(weekly_counts$start_of_week)
weekly_counts$end_of_week   <- as.Date(weekly_counts$end_of_week)

weekly_counts$days_in_week <-
  as.numeric(difftime(weekly_counts$end_of_week,
                      weekly_counts$start_of_week,
                      units = "days")) + 1

weekly_counts <- subset(weekly_counts, days_in_week == 7)

weekly_counts$CrimesScaled <-
  weekly_counts$Crimes * (7 / weekly_counts$days_in_week)

weekly_counts$Index <- seq_len(nrow(weekly_counts))

march_2020_index <- which(weekly_counts$YearWeek == "2020-09")

plot(
  weekly_counts$Index,
  weekly_counts$CrimesScaled,
  type  = "l",
  col   = "blue",
  lwd   = 2,
  xlab  = "Weekly Index",
  ylab  = "Number of Crimes (Scaled to 7 Days)"
)

abline(
  h   = pretty(weekly_counts$CrimesScaled),
  v   = pretty(weekly_counts$Index),
  col = "gray",
  lty = "dotted"
)

abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)

text(
  x     = weekly_counts$Index[march_2020_index] - 80,
  y     = 10,
  labels = "March 2020",
  pos   = 3,
  col   = "red"
)

```

## Homicides Over Time by Income Quartile

```{r, results='asis'}
#| fig.align: 'center'
library(plotly)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "
  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
  FROM (
    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
    FROM crime_la AS c
    INNER JOIN crimes_with_zips AS z 
      ON c.DR_NO = z.DR_NO 
      AND c.LAT = z.LAT 
      AND c.LON = z.LON 
      AND c.X = z.X 
      AND c.year = z.year
    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
")

homicides <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
          ref = mean(S1903_C03_001E),
          .groups = "drop")

homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
# removing NAs
homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")

homicide_trends <- homicide_trends %>%
  mutate(income_quartile = as.factor(income_quartile)) %>%
  rename("Year" = "year",
         "Homicide Count" = "homicide_count",
         "Income Quartile" = "income_quartile"
         )
# Plot homicide trends over time by income quartile
p <- ggplot(homicide_trends, aes(x = Year, y = `Homicide Count`)) +
  geom_area(aes(fill = `Income Quartile`), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)

ggplotly(p)
```

## Homicides and Median Income By Zip Code

```{r, echo=FALSE}
#| fig.align: 'center'
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT i.*, c.*
FROM cali_median_income i
INNER JOIN crimes_with_zips c
ON i.year = c.year AND CAST(i.ZCTA AS int) = c.ZCTA5CE20")
crime_income <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT `Crm.Cd.Desc`, year, DR_NO FROM crime_la")
crime_la <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)

crime_type <- crime_la %>% dplyr::select(year, Crm.Cd.Desc, DR_NO)

crime_income_type <- crime_income[,-c(249, 247)] %>% inner_join(crime_type, join_by("year" == "year", "DR_NO" == "DR_NO"))

homicide_income_by_year <- crime_income_type %>% 
  filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>%
  group_by(year, ZCTA5CE20) %>%
  summarize(HomicideCount = n(),
            cali_median_income = median(as.numeric(S1903_C03_001E))) %>%
  ungroup()

homicide_income_by_year <- homicide_income_by_year %>% filter(!is.na(HomicideCount) & !is.na(cali_median_income))

homicide_income_by_year$label <- ifelse(
  homicide_income_by_year$cali_median_income > quantile(homicide_income_by_year$cali_median_income, 0.99) | 
  homicide_income_by_year$HomicideCount > quantile(homicide_income_by_year$HomicideCount, 0.99), 
  homicide_income_by_year$ZCTA5CE20, 
  "")

ggplot(data = homicide_income_by_year, aes(x = cali_median_income, 
                                        y = HomicideCount)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, nudge_x = 2, nudge_y = 1) + 
  labs(x = "Median Annual Income ($)",
       y = "Annual Homicide Count",
       color = "Year") +
  ggthemes::theme_fivethirtyeight() + 
  theme(axis.title = element_text())
```

## Geographical Trends in Crime Rates By LAPD Bureau
```{r}
#| fig.align: 'center'
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT `AREA.NAME`, year, COUNT(*) AS numberOfCrimes FROM crime_la GROUP BY `AREA.NAME`, year")
crime_by_area <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT * FROM lapd_bureaus")
bureaus <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
bureaus$Division <- gsub("North", "N", bureaus$Division)
bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)

crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau)) +
  labs(y = "Annual Number of Crimes")
```

## Evictions and Crime Levels

```{r}
#| fig.align: 'center'
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT * FROM evictions_city_la")
evictions_la <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT `Crm.Cd.Desc`, year, month, DR_NO FROM crime_la")
crime_la <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

# convert to date
evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")

# get unique months
unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))

evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")

# create color scheme
crime_la$month_year <- paste0(crime_la$year, "-", sprintf("%02d", crime_la$month))

crime_la_monthly <- crime_la %>% 
  group_by(month_year) %>% 
  summarize(num_crimes = n())

evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")

avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)

evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), color = "white", bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_fivethirtyeight() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.y = element_text(),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```

## Geographical Distribution of Homicides
:::::: columns
:::: {.column width="35%"}
::: {style="font-size: 50%"}
```{r}
sliderInput("selected_year", "Select Year:",
                  min = 2010, max = 2024,
                  value = 2010, sep = "", step= 1)
```
:::
::::

::: {.column width="65%"}
```{r}
library(leaflet)
leafletOutput("crime_map", height = 600)
```
:::
::::::

```{r}
#| context: server
library(leaflet)
library(RSQLite)
library(shiny)
library(leaflet)
library(DBI)
library(RSQLite)
library(dplyr)

crime_data <- reactive({
    req(input$selected_year)
    con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
    crime_la_coords <- dbGetQuery(con, paste("SELECT LAT, LON, year FROM crime_la WHERE `Crm.Cd.Desc` = 'CRIMINAL HOMICIDE' AND year = ", input$selected_year))
    dbDisconnect(con)
    return(crime_la_coords)
  })

output$crime_map <- renderLeaflet({ 
  filtered <- crime_data()
  leaflet() %>%
  addTiles(
    urlTemplate = "https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png",
    attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/">CARTO</a>'
  ) %>%
  setView(lat = 34.003, lng = -118.3, zoom = 10) %>% 
  addCircleMarkers(data=filtered, lng=~LON, lat=~LAT)
  })
```

# Correlations and Modeling {background="#6B6B6B"}

## Correlation Heat Map
```{r}
#| fig.align: 'center'
library(lubridate)
# creating dataset for correlations
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions.
# covid data
query <- dbSendQuery(con, "
  SELECT 
    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
   cases_14day
  FROM covid_cases
")
covid_14day <- dbFetch(query,n=-1)
dbClearResult(query)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_14day$date + 7

covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)

covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)

covid_la_monthly$index <- 1:nrow(covid_la_monthly)

# next, unemployment rate, lapd budget, and evictions
query <- dbSendQuery(con, "
  SELECT 
    u.month, 
    u.Year, 
    u.Value AS unemployment_rate, 
    l.police_budget, 
    e.Count AS num_evictions
  FROM 
    unemployment_la AS u
  INNER JOIN 
    lapd_budget AS l ON u.Year = l.year
  INNER JOIN 
    evictions_county_la AS e ON u.Year = e.Year AND u.month = e.Month
")
unemployment_lapd_budget_evictions <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)

crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
crime_la_monthly$month <- month(crime_la_monthly$date)
crime_la_monthly$year <- year(crime_la_monthly$date)

correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget)

correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))

correlations <- correlations_df2 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)

Cor_mat <- round(cor(correlations), 2)

# get only the upper triangle of variables 
# (because a correlation heat map shows each correlation value twice)
get_upper_tri <- function(cormat){
  # lower.tri -> Returns a matrix of logicals, so set the values that are true to NA
  cormat[lower.tri(cormat)]<- NA # set the correlation matrix lower triangle to NA
  return(cormat) 
}

# reorder them by the strength of the correlation
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd) # use hierarchical clustering to reorder
  cormat <-cormat[hc$order, hc$order]
}

# use the functions above to remove the unnecessary parts
Cor_mat<-reorder_cormat(Cor_mat)

Cor_mat_upper<-get_upper_tri(Cor_mat)

# using melt instead of pivot_longer because pivot_longer did not order correctly
Cor_mat_upper<-reshape2::melt(Cor_mat_upper, na.rm=TRUE)

# remove the central values 
# which show the correlation for each variable with itself
Cor_mat_upper<- Cor_mat_upper %>% filter(value != 1.00)

long_Cor <- Cor_mat_upper %>%
    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```

## Relationship Between Homicides and Socioeconomic Variables
```{r}
#| context: setup
#| include: false
library(RSQLite)
library(ggplot2)
library(dplyr)

corr <- read.csv("corr.csv")

colnames(corr) <- c("X", "Year", "Month", "Unemployment", "Crimes", "COVID", "Evictions", "LAPD Budget")
```

```{r}
#| panel: sidebar
radioButtons("opts", "View homicides compared to:", 
            choices = c("COVID", "Evictions", "LAPD Budget", "Unemployment"))
```

```{r}
#| panel: fill
plotOutput("scatter")
```

```{r}
#| context: server
output$scatter <- renderPlot({
  # Guessing 'value' is the comparison variable — adjust as needed
  df <- corr[c("Crimes", input$opts, 'Year')]
  colnames(df) <- c("Crimes", "var", "Year")

  ggplot(df, aes(x = var, y = Crimes)) +
    geom_point(aes(colour = as.factor(Year))) +
    scale_color_viridis_d() +
  ggthemes::theme_fivethirtyeight() +
  theme(axis.title = element_text()) + 
    labs(x = input$opts, y = "Crime Count", title = paste("Monthly Crime vs", input$opts),
         colour = "Year") +
    geom_smooth(method = "lm")
}) 
```

## Modeling Crime Rates Spatially

\begin{align}
\text{HomicideCount}_i = \beta_0 + 
\beta_1 \text{Latitude}_i +
\beta_2 \text{Longitude}_i + \epsilon_i
\end{align}

Homicide rates are higher in the southeastern parts of Los Angeles.

```{r, warning=FALSE, message=FALSE, fig.align = "center"}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "
  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
  FROM (
    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
    FROM crime_la AS c
    INNER JOIN crimes_with_zips AS z 
      ON c.DR_NO = z.DR_NO 
      AND c.LAT = z.LAT 
      AND c.LON = z.LON 
      AND c.X = z.X 
      AND c.year = z.year
    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
")

homicides <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

homicides_latlon_rounded <- homicides %>%
  mutate(LAT_rounded = round(LAT, 2),
         LON_rounded = round(LON, 2))

homicides_lat_long_yr <- homicides_latlon_rounded %>% 
  group_by(year, LAT_rounded, LON_rounded) %>%
  summarize(HomicideCount = n()) %>% ungroup()

linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)

d <- coef(summary(linear_model1))[,c(1, 4)]
colnames(d) <- c("Coefficient", "p-value")
rownames(d) <- c("(Intercept", "Latitude", "Longitude")
knitr::kable(d)
```

## Homicides Increased In Pandemic Years

\small
\begin{equation}
\text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
\end{equation}

```{r}
#| fig.align: 'center'
#| fig.width: 6
#| fig.height: 4
avg <- homicides_lat_long_yr %>% group_by(year) %>% summarize(avgHC = mean(HomicideCount)) %>% ungroup()
ggplot(avg, aes(x = year, y = avgHC)) +
  geom_line(color = "#7e03a8") +
  ggthemes::theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        plot.caption = element_text(size = 8)) +
  labs(y = "Average Homicide Count",
       x = "Year",
       title = "Average Homicide Count By Year",
       caption = "Average was derived from latitude and longitude groups used in the model.")
```

# Killer Plot {background="#6B6B6B"}

## Walk of Crime

```{r}
#| context: setup
#| include: false
# load functions
source("dependencies.R")
library(dplyr)
spatial <- read.csv("homicides.csv")
# load data
con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))

crime_la_annually <- dbGetQuery(con, "SELECT * FROM crime_la_annually")
covid_la_annually <- dbGetQuery(con, "SELECT * FROM covid_la_annually")
evictions_county_la <- dbGetQuery(con, "SELECT * FROM evictions_county_la")
lapd_budget_annually <- dbGetQuery(con, "SELECT * FROM lapd_budget_annually")
unemployment_la <- dbGetQuery(con, "SELECT * FROM unemployment_la")
correlations_summary <- dbGetQuery(con, "SELECT * FROM correlations_quarterly")
correlations_summary$period <- 1:nrow(correlations_summary)

map_datasets <- list("Crime Count" = crime_la_annually,
                     "Covid Cases" = covid_la_annually,
                     "Evictions" = evictions_county_la,
                     "LAPD Budget" = lapd_budget_annually,
                     "Unemployment Rate" = unemployment_la)

dbDisconnect(con)
```

```{r}
#| panel: sidebar

radioButtons("type", "View:", choices = c("Homicides Spatially", "A Variable Over Time"), selected = "Homicides Spatially")
# Dropdown to select dataset
uiOutput("dataset")

uiOutput("year_selector")

uiOutput("crime_type_ui")
```

```{r}
#| panel: fill

plotOutput("stars")
```

```{r}
#| context: server

selectedKillerData <- reactive({
  req(input$dataset, input$year_range)

  if (input$dataset == 'Crime Count') {
    req(input$crime_type)
    df <- map_datasets[[input$dataset]] %>%
      filter(year %in% seq(input$year_range[1], input$year_range[2]),
             Crm.Cd.Desc == input$crime_type)
  } else {
    df <- map_datasets[[input$dataset]] %>%
      filter(year %in% seq(input$year_range[1], input$year_range[2]))
  }
  return(df)
})

output$dataset <- renderUI({
  req(input$type)

  if (input$type == "A Variable Over Time") {
  selectInput("dataset", "Choose Dataset", 
            choices = c("Crime Count", 
                        "Covid Cases", 
                        "Evictions", 
                        "LAPD Budget", 
                        "Unemployment Rate")) 
}
})

output$year_selector <- renderUI({
  req(input$type)

  if (input$type == "Homicides Spatially") {
    # Show this when Homicides Spatially is selected
    selectInput("year_single", "Select Year:", choices = 2019:2023, selected = 2023)
  } else {
    req(input$dataset)  # Now only required when viewing variables over time

    # Get relevant years
    year_vals <- switch(input$dataset,
                        "Crime Count" = crime_la_annually$year,
                        "Covid Cases" = covid_la_annually$year,
                        "Evictions" = evictions_county_la$year,
                        "LAPD Budget" = lapd_budget_annually$year,
                        "Unemployment Rate" = unemployment_la$year)

    year_vals <- sort(unique(year_vals))

    sliderInput("year_range", "Select Time Range:",
                min = min(year_vals), max = max(year_vals),
                value = c(min(year_vals), max(year_vals)), step = 1,
                sep = "")
  }
})

output$crime_type_ui <- renderUI({
  req(input$type)
  req(input$dataset)

  if (input$type == "A Variable Over Time" && input$dataset == "Crime Count") {
    selectInput("crime_type", "Select Crime Type",
                choices = unique(crime_la_annually$Crm.Cd.Desc),
                selected = unique(crime_la_annually$Crm.Cd.Desc)[1])
  }
})

output$stars <- renderPlot({
  req(input$type)

  if (input$type == "Homicides Spatially") {
    req(input$year_single)
    print(as.integer(input$year_single))
    print(unique(spatial$year))
    df <- spatial %>% filter(year == as.integer(input$year_single))
    df$LAT <- round(df$LAT_rounded, 1)
    df$LON <- round(df$LON_rounded, 1)

    df <- df %>% group_by(LAT, LON) %>% summarize(avgHomicides = mean(HomicideCount)) %>% ungroup()

    df$LATstd <- .175 + ((df$LAT - min(df$LAT))/(max(df$LAT) - min(df$LAT))) * .75
    df$LONstd <- .175 + ((df$LON - min(df$LON))/(max(df$LON) - min(df$LON))) * .75
   labels <- levels(standardize_scale(df$avgHomicides))
   df$scale <- as.numeric(standardize_scale(df$avgHomicides))

   draw_spatial_killer(df, labels)
} else {
      df <- selectedKillerData()
      rows <- nrow(df)
      col_row <- if (rows <= 13) {
        c(rows, 1)
      } else if (rows == 14) {
        c(7, 2)
      } else if (rows == 15) {
        c(5, 3)
      } else if (rows == 16) {
        c(4, 4)
      } else {
        c(ceiling(sqrt(rows)), ceiling(rows / ceiling(sqrt(rows))))
      }

      stars <- generate_star_centers(col_row[1], col_row[2])
      labels <- levels(standardize_scale(df$variable))
      df_sorted <- df[order(df$year), ]

      stars$scale <- df_sorted$bins
      stars$year <- df_sorted$year

      draw_timeline_killer(stars, labels, input$dataset)
       }
    })
```